DATA 624 Homework 8

### 2020-04-14 DATA624/HW8.Rmd
More work on HW 8

```{r}
library(AppliedPredictiveModeling)
data(permeability)
df <- as.data.frame(fingerprints[, nearZeroVar(fingerprints)]) %>%
  mutate(y = permeability)

set.seed(42)

in_train <- createDataPartition(df$y, times = 1, p = 0.8, list = FALSE)
train_df <- df[in_train, ]
test_df <- df[-in_train, ]

pls_model <- train(
  y ~ ., data = train_df, method = "pls",
  center = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)

pls_model
```

```{r}
pls_predictions <- predict(pls_model, test_df)

results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$y))) %>%
  mutate("Model"= "PLS")
```

@@

```{r}
knn_model <- train(y ~ ., data = train_df, method = "knn",
                   preProcess = c("center"), 
                   trControl = trainControl("cv", number = 10),
                   tuneLength = 25)
knn_model

@@

results <- data.frame(t(postResample(pred = knn_predictions, obs = test_df$y))) %>%
  mutate("Model"= "KNN") %>% rbind(results)
```

@@

```{r}
MARS_grid <- expand.grid(.degree = 1:2, .nprune = 2:15)
MARS_model <- train(y ~ ., data = train_df, method = "earth",
                    tuneGrid = MARS_grid,
                    preProcess = c("center"),
                    tuneLength = 25)
MARS_model

@@

results <- data.frame(t(postResample(pred = MARS_predictions, obs = test_df$y))) %>%
  mutate("Model"= "MARS") %>% rbind(results)
```

@@

```{r}
SVM_model <- train(y ~ ., data = train_df, method = "svmRadial",
                   preProcess = c("center"),
                   tuneLength = 25,
                   trControl = trainControl(method = "cv"))
SVM_model

@@

results <- data.frame(t(postResample(pred = SVM_predictions, obs = test_df$y))) %>%
  mutate("Model"= "SVM") %>% rbind(results)
```

@@

@@

@@

Explore the relationships between the top predictors and the response for the predictors that are unique to the optimal nonlinear regression model. Do these plots reveal intuition about the biological or process predictors and theri relationship with yield?

>>>>

**I will run the data through the models in the chapter trying to keep each model as close to the original linear model in terms of the parameters.**

```{r}
library(AppliedPredictiveModeling)
data(ChemicalManufacturingProcess)

# Make this reproducible
set.seed(42)

knn_model <- preProcess(ChemicalManufacturingProcess, "knnImpute")
df <- predict(knn_model, ChemicalManufacturingProcess)

df <- df %>%
  select_at(vars(-one_of(nearZeroVar(., names = TRUE))))

in_train <- createDataPartition(df$Yield, times = 1, p = 0.8, list = FALSE)
train_df <- df[in_train, ]
test_df <- df[-in_train, ]

pls_model <- train(
  Yield ~ ., data = train_df, method = "pls",
  center = TRUE,
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)

pls_model

pls_predictions <- predict(pls_model, test_df)

results <- data.frame(t(postResample(pred = pls_predictions, obs = test_df$Yield))) %>%
  mutate("Model"= "PLS")
```

@@

```{r}
knn_model <- train(
  Yield ~ ., data = train_df, method = "knn",
  center = TRUE,
  scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)
knn_model

@@

results <- data.frame(t(postResample(pred = knn_predictions, obs = test_df$Yield))) %>%
  mutate("Model"= "KNN") %>% rbind(results)
```

@@

```{r}
MARS_grid <- expand.grid(.degree = 1:2, .nprune = 2:15)

MARS_model <- train(
  Yield ~ ., data = train_df, method = "earth",
  tuneGrid = MARS_grid,
  # If the following lines are uncommented, it throws an error
  #center = TRUE,
  #scale = TRUE,
  trControl = trainControl("cv", number = 10),
  tuneLength = 25
)
MARS_model

@@

results <- data.frame(t(postResample(pred = MARS_predictions, obs = test_df$Yield))) %>%
  mutate("Model"= "MARS") %>% rbind(results)
```

@@

```{r}
SVM_model <- train(
  Yield ~ ., data = train_df, method = "svmRadial",
  center = TRUE,
  scale = TRUE,
  trControl = trainControl(method = "cv"),
  tuneLength = 25
)
SVM_model

@@

results <- data.frame(t(postResample(pred = SVM_predictions, obs = test_df$Yield))) %>%
  mutate("Model"= "SVM") %>% rbind(results)
```

#### Neural Network Model

```{r}
nnet_grid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)
nnet_maxnwts <- 5 * ncol(train_df) + 5 + 1
nnet_model <- train(
  Yield ~ ., data = train_df, method = "avNNet",
  center = TRUE,
  scale = TRUE,
  tuneGrid = nnet_grid,
  trControl = trainControl(method = "cv"),
  linout = TRUE,
  trace = FALSE,
  MaxNWts = nnet_maxnwts,
  maxit = 500
)

nnet_model

nnet_predictions <- predict(nnet_model, test_df)

results <- data.frame(t(postResample(pred = nnet_predictions, obs = test_df$Yield))) %>%
  mutate("Model"= "Neural Network") %>% rbind(results)

@@

**The SVM Model was the best non-linear model.**

@@

```{r}
varImp(SVM_model, 10)
```

```{r}
varImp(pls_model, 10)
```

**The SVM model was very similar to the PLS model.  In either case the manufacuturing process variables dominate the list. The SVM model found `BiologicalMaterial12` to be important and didn't find `BiologicalMaterial02` to be important.**

@@

Explore the relationships between the top predictors and the response for the predictors that are unique to the optimal nonlinear regression model. Do these plots reveal intuition about the biological or process predictors and their relationship with yield?

```{r}
ggplot(train_df, aes(BiologicalMaterial12, Yield)) +
  geom_point()
```

**This indicates a positive relationship between the `BiologicalMaterial02` and the yeild, although it seems that there is a sweet spot, or a point of diminishing marginal returns on the material as the highest yeilds are not the furthest to the right on the graph.**
